---
title: "goJitsu"
author: "Cade Michael Lueker"
date: "12/29/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Goal and Methods 

* After taking class in R and basic data analytics I thought that it would be fun
to try and find / scrape a data set of my own. Not one that was given or loaded from
elsewhere. I decided that I would try to get my hands on some Jiu Jitsu data and 
turned to **BJJHeroes** to do so.
* as of right now my data is comprised of a 
  * name 
  * number of wins
  * number of losses
  * number per type of win
  * number per type of loss
* this data is relatively uninteresting as it is fairly simple and doesn't 
allow for much prediction, ie predicting if one fighter would beat another. 
* But scraping the data was a foreign task and I will try to get more detailed
data in the near future to update and or re write this report. 

#### Data Collection

* to collect the data I decided to try **Go** as it is a language that I am very interested in 
and want to become more familiar with. 
  * I utilized the go colly library and found that it offered a very clean and consise meta 
  for web scraping.
  * I visit a page that lists fighters and every fighter's individual page gets 
  visited for individual data collection
  * if this fighter has wins or losses (some fighters only have descriptions) their 
  unique data gets entered into a CSV engendered by the program. 
*the go code is at the end of this report* 

#### Possible observations

* The only real observations I can make with the data I collected is to see the distribution
of different win & loss types. I can compare the frequencies of each, I can determine what 
a median and or average competitor (who is well known enough for BJJ heroes) looks like 
in terms of wins and losses.
* I will rely mainly on histograms as they are utilized to measure frequency which is what I can 
analyze.
* I will use basic linear models to see if certain win type frequencies can predict more wins overall.

## Load Required packages

```{r}
# load libraries
library(tidyverse)
library(ggplot2)
library(vioplot)
```

## Read The data from the CSV 

```{r}
goData <- read.csv("./workingFiles/winTypes.csv")
```

## Clean data

```{r}
goData$wins <- gsub("[A-Z]", "", goData$wins)
goData$losses <- gsub("[A-Z]", "", goData$losses)
nonNumer <- names(goData)[2:length(goData)]
goData[nonNumer] <- lapply(goData[nonNumer], as.numeric)
# add total number of matches 
goData <- goData %>%
  mutate(total = losses + wins)
# convert to R df 
winLoss_df <- as.data.frame(goData)

# add number of matches won not by submission 
winLoss_df <- winLoss_df %>%
  mutate(nonSubWins = wins - Win.Submissions)

# separate  wins 
wins_df <- winLoss_df %>% select(-c(colnames(winLoss_df[11:17]),"nonSubWins"))
# calculate submission percentage for wins 
wins_df <- wins_df %>%
  mutate(subRate = Win.Submissions / wins) %>% # change between total and wins
  mutate(winRate = wins / total) %>% 
  mutate(pWinRate = Win.Points / wins)
wins_df[is.na(wins_df)] <- 0
```

## Initial Plots on match data, Histograms

```{r}
# Basic Histograms 
hist(winLoss_df$Losses.Points,
     ylim = range(0,160),
     xlab = "Number of Point Losses",
     main = "Frequency of Point Losses",
     breaks = 25)
hist(winLoss_df$Losses.Submissions, 
     ylim = range(0,160),
     xlab = "Number of Submission Losses", 
     main = "Frequency of Submission Losses",
     breaks = 25)
hist(winLoss_df$Win.Points,
     ylim = range(0,160),
     xlab = "Number of Point Wins",
     main = "Frequency of Point Wins",
     breaks = 25)
hist(winLoss_df$Win.Submissions, 
     ylim = range(0,160),
     xlab = "Number of Submission Wins", 
     main = "Frequency of Submission Wins",
     breaks = 25)
# number of losses for Subs and Points 
sum(winLoss_df$Losses.Submissions)
sum(winLoss_df$Losses.Points)
```

## Box Plots For wins

```{r}
# sub wins vs points wins 
winBar <- winLoss_df %>%
  ggplot() + 
    geom_histogram(aes(x = Win.Points,fill= "blue"),
                   color = "black", alpha = 0.75) +
    geom_vline(aes(xintercept = mean(Win.Points)),
               color = "dark blue", linetype = "dashed") + 
    geom_histogram(aes(x = Win.Submissions, fill= "green"),
                   color = "black", alpha = 0.5) + 
    geom_vline(aes(xintercept = mean(Win.Submissions)),
               color = "dark green", linetype = "dashed") +
    labs(x = ("Number of Wins"),
         y = ("Frequency"),
         title = "Submissions vs Point Wins") +
    scale_fill_identity(name = "fill color",
                        guide = "legend",
                        labels = c("Pts", "Subs")) +
    theme_classic()
# plot 
winBar

# sub wins vs all other types of wins 
subVsNon <- winLoss_df %>%
  ggplot() + 
    geom_histogram(aes(x = nonSubWins, fill= "purple"),
                   color = "black", alpha = 1) +
    geom_vline(aes(xintercept = mean(nonSubWins)),
               color = "purple", linetype = "dashed") + 
    geom_histogram(aes(x = Win.Submissions, fill= "green"),
                   color = "black", alpha = 0.5) + 
    geom_vline(aes(xintercept = mean(Win.Submissions)),
               color = "dark green", linetype = "dashed") +
    labs(x = ("Number of Wins"),
         y = ("Frequency"),
         title = "Submissions vs All Other Wins") +
    scale_fill_identity(name = "fill color",
                        guide = "legend",
                        labels = c("Subs", "NonSubs")) +
    theme_classic()

# plot
subVsNon

# wins by points vs all other types of wins 
PtsVsNon <- winLoss_df %>%
  ggplot() + 
    geom_histogram(aes(x = wins - Win.Points, fill= "purple"),
                   color = "black", alpha = 1) +
    geom_histogram(aes(x = Win.Points, fill= "blue"),
                   color = "black", alpha = 0.5) + 
    labs(x = ("Number of Wins"), y = ("Frequency"),
         title = "Points vs All Other Wins") +
    scale_fill_identity(name = "fill color",
                        guide = "legend",
                        labels = c("Pts", "NonPts")) +
    theme_classic()

# plot
PtsVsNon
```

### Violin Plots

```{r}
winViolin <- vioplot(winLoss_df$wins,
                         col = "purple",
                         names = "Frequency",
                         ylab = "Number of Wins")

winViolin

subViolin <- vioplot(winLoss_df$Win.Submissions,
                         col = "green",
                         names = "Frequency",
                         ylab = "Number of Subs")

subViolin

allTogether <- vioplot(
  winLoss_df$wins,
  winLoss_df$Win.Points,
  winLoss_df$Win.Submissions,
  col = c("purple", "blue", "green"),
  names = c("Total Wins", "Points", "Subs"),
  ylab = "Number of Wins",
  xlab = "Frequency"
)
```

## Finishing rate (of Wins) & Win Rate (of total matches)

```{r}
# keep in mind this is submissions / total wins  
hist(wins_df$subRate * 100, 
     xlab = "Finishing % (per match)",
     main = "Frequency of Submission Rate",
     breaks = 25)
# violin plot of subrate 
subRateViolin <- vioplot(wins_df$subRate * 100,
                         col = "green",
                         names = "Frequency",
                         ylab = "SubMission Rate")

# plot the density of the submission rate 
plot(density(wins_df$subRate * 100), main = "Density of Submission Rate")

# Win Rate 
hist(wins_df$winRate * 100, 
     xlab = "Win % (per match)",
     main = "Frequency of Win Rate",
     breaks = 25)
# violin plot of winRate 
subRateViolin <- vioplot(wins_df$winRate * 100,
                         col = "purple",
                         names = "Frequency",
                         ylab = "Win Rate")

# plot the density of the Win rate 
plot(density(wins_df$winRate * 100), main = "Density of Win Rate")

# quick look at point win rate 

plot(density(wins_df$pWinRate * 100), main = "Density of Point Win Rate")

# comparing win rate and submission rate on a graph with a quick graph 
plot(wins_df$winRate * 100, wins_df$subRate * 100,
     xlab = "Win Rate",
     ylab = "Sub Rate",
     main = "Win Rate vs Sub Rate",
     col = "green",
     pch = 16,
     cex = 0.75)
```


## basic modeling to predict Win Rate 


### Subs Rate (of wins) vs Win Rate (of total matches)

```{r}
winVsSub <- wins_df %>%
  ggplot(aes(y= subRate, x = winRate)) +
  geom_point(color = "green") +
  labs(title = "Win Rate vs Sub Rate", y = "Sub Rate", x = "Win Rate") +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE , color = "black") +
  theme(legend.position="none") +
  theme_classic()

winVsSub

# turn the line from the above graph into a model
simpleModel <- lm(wins_df$subRate ~ wins_df$winRate, wins_df)
# show the summary stats of the model
summary(simpleModel)

# plot the residuals of the model (want to see symmetry)
residualsPlot <- simpleModel %>%
  ggplot(aes(x = .fitted, y = .resid)) +
  geom_point(color = "green") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
  xlab("Fitted values") +
  ylab("Residuals") +
  theme_classic()
# show the plot
residualsPlot

# graphing the residuals
qqnorm(simpleModel$residuals,
  ylab="Standardized Residuals",
  xlab="Normal Scores")
qqline(simpleModel$residuals)

# check the distribution of the residuals (want normality)
hist(simpleModel$residuals, main = "Histogram of Model Residuals (S win)", xlab = "Residuals")
```

### Point Rate (of wins) vs Win Rate (of total matches)

```{r}
winVsSub <- wins_df %>%
  ggplot(aes(y= pWinRate, x = winRate)) +
  geom_point(color = "blue") +
  labs(title = "Win Rate vs Point Win Rate", y = "P Win Rate", x = "Win Rate") +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE , color = "black") +
  theme(legend.position="none") +
  theme_classic()

winVsSub

# turn the line from the above graph into a model
simpleModel2 <- lm(wins_df$pWinRate ~ wins_df$winRate, wins_df)
# show the summary stats of the model
summary(simpleModel2)

# plot the residuals of the model (want to see symmetry)
residualsPlot2 <- simpleModel2 %>%
  ggplot(aes(x = .fitted, y = .resid)) +
  geom_point(color = "blue") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
  xlab("Fitted values") +
  ylab("Residuals") +
  theme_classic()
# show the plot
residualsPlot2

# graphing the residuals
qqnorm(simpleModel2$residuals,
  ylab="Standardized Residuals",
  xlab="Normal Scores")
qqline(simpleModel2$residuals)

# check the distribution of the residuals (want normality)
hist(simpleModel2$residuals, main = "Histogram of Model Residuals (P win)", xlab = "Residuals")
```

## Conclusion

* the Medians for submissions and total wins are 13, and 32 respectively. This means that to be 
in the upper $50\%$ of competitors you need to have 32 wins or more and to be in the upper range of 
submission artists you need $13$ or more submissions. 
* initially when running the models I had submission rate and point win rate of
all the matches, but this was almost a tautology. Of course someone more likely 
to win a match by submission or points than not win will have a high win rate overall.
So I changed the models to look at sub / point win rate of total wins. This way someone 
could have an extremely high submission rate but a very low win rate with less variable 
dependence. 
* Based on the models $R^2$ values we can see a very slight **positive correlation** between
the likelihood that a match won by a competitor was won by a submission and the likelihood that
that same competitor would win a match via any method. Inversely we can see a very slight **negative correlation** between a competitors likelihood to have won a match by points and their overall likelihood to 
win a match. 
* because the p value of the model looking at submission rate is $0.0002024$ we can say that 
winning more matches by submission than by not submission does signify a greater chance to win a match.
* however, because the p value of the model looking at is above $0.05$ we cannot say that having a 
high likelihood of wins coming from points makes you more likely to lose a match. 
* in summary this shows that **more submissions == better Jiu Jitsu Competitor!!!** (but we already knew that)


## Go code for scraping

```go
package main

import (
	"encoding/csv"
	"fmt"
	"log"
	"os"

	"github.com/gocolly/colly"
)

// main function  
func main() {
	createList()
	visitAll()
	indiv("https://www.bjjheroes.com/?p=7556")
}

// visitAll function  
func visitAll() {
	/* instatiate colly */
	c := colly.NewCollector(
		colly.AllowedDomains(
			"www.bjjheroes.com/",
			"bjjheroes.com/",
			"https://bjjheroes.com/",
			"www.bjjheroes.com",
			"bjjheroes.com",
			"https://bjjheroes.com",
		),
	)
	// if a link is found in a table
	c.OnHTML("td.column-1 a[href]", func(e *colly.HTMLElement) {
		url := e.Request.AbsoluteURL(e.Attr("href")) // get the link as an absolute
		indiv(url)                                   // run indiv on the absolute link
	})
	// start scraping
	c.Visit("https://www.bjjheroes.com/a-z-bjj-fighters-list")
}

// createList function  
func createList() {
	// instantiate csv
	file, err := os.OpenFile("list.csv", os.O_APPEND|os.O_CREATE|os.O_WRONLY, 0644)
	if err != nil { // deal with errors
		log.Fatalf("cannot create file due to %s", err)
		return
	}
	defer file.Close()
	writer := csv.NewWriter(file) // create a writer to modify csv
	// write column names to csv
	writer.Write([]string{
		"Name",
		"wins",
		"losses",
		"Win Points",
		"Win Advantages",
		"Win Submissions",
		"Win Decision",
		"Win Penalties",
		"Win EBI / OT",
		"Win DQ",
		"Losses Points",
		"Losses Advantages",
		"Losses Submissions",
		"Losses Decision",
		"Losses Penalties",
		"Losses EBI / OT",
		"Losses DQ",
	})
	writer.Flush() // ensure data is written
}

// indiv function  
func indiv(url string) {
	// instantiate or open csv
	file, err := os.OpenFile("list.csv", os.O_APPEND|os.O_CREATE|os.O_WRONLY, 0644)
	if err != nil {
		log.Fatalf("cannot create file due to %s", err)
		return
	}
	defer file.Close()
	writer := csv.NewWriter(file) // create a writer to modify csv
	defer writer.Flush()          // defer pushes to end of function
	/* instatiate colly */
	c := colly.NewCollector(
		colly.AllowedDomains(
			"www.bjjheroes.com/",
			"bjjheroes.com/",
			"https://bjjheroes.com/",
			"www.bjjheroes.com",
			"bjjheroes.com",
			"https://bjjheroes.com",
		),
	)
	// create empty column variables
	// using standard vs short declaration as short is a bit messy in this case
	fName := "0"
	wins := "0"
	lose := "0"
	wPoints := "0"
	wAdvantage := "0"
	wSubs := "0"
	wDec := "0"
	wPen := "0"
	wEBI := "0"
	wDQ := "0"
	lPoints := "0"
	lAdvantage := "0"
	lSubs := "0"
	lDec := "0"
	lPen := "0"
	lEBI := "0"
	lDQ := "0"

  // find name 
	c.OnHTML("h1", func(a *colly.HTMLElement) {
		fName = a.Text
	})
  //find wins 
	c.OnHTML("div.Win_title", func(b *colly.HTMLElement) {
		wins = b.ChildText("em")
	})
  // find win types
	c.OnHTML("div.wrapper_canvas li", func(d *colly.HTMLElement) {
		switch d.ChildText("span.by_points") {
		case "BY POINTS":
			wPoints = d.ChildText("span.per_no")
		case "BY ADVANTAGES":
			wAdvantage = d.ChildText("span.per_no")
		case "BY SUBMISSION":
			wSubs = d.ChildText("span.per_no")
		case "BY DECISION":
			wDec = d.ChildText("span.per_no")
		case "BY PENALTIES":
			wPen = d.ChildText("span.per_no")
		case "BY EBI/OT":
			wEBI = d.ChildText("span.per_no")
		case "BY DQ":
			wDQ = d.ChildText("span.per_no")
		}
	})
  // find losses 
  c.OnHTML("div.Win_title_lose", func(t *colly.HTMLElement) {
    lose = t.ChildText("em")
  })
  // find loss types 
	c.OnHTML("div.wrapper_canvas_lose li", func(d *colly.HTMLElement) {
		switch d.ChildText("span.by_points") {
		case "BY POINTS":
			lPoints = d.ChildText("span.per_no_lose")
		case "BY ADVANTAGES":
			lAdvantage = d.ChildText("span.per_no_lose")
		case "BY SUBMISSION":
			lSubs = d.ChildText("span.per_no_lose")
		case "BY DECISION":
			lDec = d.ChildText("span.per_no_lose")
		case "BY PENALTIES":
			lPen = d.ChildText("span.per_no_lose")
		case "BY EBI/OT":
			lEBI = d.ChildText("span.per_no_lose")
		case "BY DQ":
			lDQ = d.ChildText("span.per_no_lose")
		}
	})

  // start the scraper
	c.Visit(url)

  // write to csv 
	if wins != "0" || lose != "0" {
		writer.Write([]string{
			fName,
			wins,
			lose,
			wPoints,
			wAdvantage,
			wSubs,
			wDec,
			wPen,
			wEBI,
			wDQ,
			lPoints,
			lAdvantage,
			lSubs,
			lDec,
			lPen,
			lEBI,
			lDQ,
		})
	}
  // for every individual fighter print done
	fmt.Printf("!!!Done!!!")
}
```




